---
title: "Linear Regression_ML"
author: "Joe Marco"
date: "September 26, 2017"
output: html_document
---

## Exercise: Least Squares Regression
```{r}
suppressMessages(library(ggplot2))
suppressMessages(library(tidyverse))
setwd("~/DataScienceFoundationsCourse/MachineLearning/LinearRegression")
```


### Use the /states.rds/ data set. Fit a model predicting energy consumed per capita (energy) from the percentage of residents living in metropolitan areas (metro).

```{r}
# read in states file and assign it to the "states" dataframe
states <- readRDS("states.rds")

#inspect the structure of states
str(states)
```

* 1. Examine/plot the data before fitting the model
```{r}
# create dataframe with the subset of metro and energy fields
states_en_met <- subset(states, select = c("metro", "energy"))

#inspect the structure of the new data frame
str(states_en_met)

#inspect further a summary of the new data frame
summary(states_en_met)

# check the correlation
cor(states_en_met, use="pairwise")

#plot the data frame, as we can see from the below, appears to be a negative correlation hence the best line we could fit against the plot would show that as metro increases, energy levels decrease
plot(states_en_met)
```

* 2. Print and interpret the model `summary'

```{r}
#create a linear model of the assigned data and assign it to model_en_met
model_en_met <- lm(energy ~ metro, data = states)

#view the summary of our model. We can see that "metro" has significance in our model as defined by the * next to our value
summary(model_en_met)
```

* 3. `plot' the model to look for deviations from modeling assumptions

```{r}
plot(model_en_met)
```

* Select one or more additional predictors to add to your model and repeat steps 1-3. Is this model significantly better than the model with /metro/ as the only predictor?
```{r}
states_en_met_pop <- subset(states, select = c("energy", "metro", "pop"))
summary(states_en_met_pop)
plot(states_en_met_pop)

# check correlation
cor(states_en_met_pop, use="pairwise")

## new model
model_en_met_pop <- lm(energy ~ metro + pop + waste, data = states)
summary(model_en_met_pop)
```


## Exercise: interactions and factors

*Use the states data set.

* 1. Add on to the regression equation that you created in exercise 1 by generating an interaction term and testing the interaction.

```{r}
mod_en_metro_by_waste <- lm(energy ~ metro * waste, data = states)
```


* 2. Try adding region to the model. Are there significant differences across the four regions?

```{r}
mod_en_region <- lm(energy ~ metro * waste + region, data = states)
anova(mod_en_region)
```

