---
title: "test"
author: "Joe Marco"
date: "October 4, 2017"
output: html_document
---
```{r}
data(wine, package="rattle.data")
head(wine)
```

data(wine, package="rattle.data")
head(wine)

### Exercise 1: Remove the first column from the data and scale it using the scale() function:
```{r}
wine$Type <- NULL

#double check column has been removed properly
head(wine)

#apply scale function
wine <- scale(wine)
```


#### Now we'd like to cluster the data using K-Means. How do we decide how many clusters to use if you don't know that already? We'll try two methods.

#### Method 1: A plot of the total within-groups sums of squares against the number of clusters in a K-means solution can be helpful. A bend in the graph can suggest the appropriate number of clusters. 
```{r}
wssplot <- function(data, nc=15, seed=1234){
  wss <- (nrow(data)-1)*sum(apply(data,2,var))
  for (i in 2:nc){
    set.seed(seed)
    wss[i] <- sum(kmeans(data, centers=i)$withinss)}
  
  plot(1:nc, wss, type="b", xlab="Number of Clusters",
       ylab="Within groups sum of squares")
}

wssplot(wine)
```
* How many clusters does this method suggest?
* Why does this method work? What's the intuition behind it?
* Look at the code for wssplot() and figure out how it works


### Exercise 2:

#### Method 2: Use the NbClust library, which runs many experiments and gives a distribution of potential number of clusters.
```{r}
distances = dist(wine, method = "euclidean")
clusterwine = hclust(distances, method = "ward")
plot(clusterwine)
```


```{r}
library(NbClust)
set.seed(1234)
nc <- NbClust(wine, min.nc=2, max.nc=15, method="kmeans")
table(nc$Best.n[1,])
```

```{r}
barplot(table(nc$Best.n[1,]),
        xlab="Numer of Clusters", ylab="Number of Criteria",
        main="Number of Clusters Chosen by 26 Criteria")
```


#### How many clusters does this method suggest?
* 3 clusters



### Exercise 4: Once you've picked the number of clusters, run k-means using this number of clusters. Output the result of calling kmeans() into a variable fit.km
```{r}
set.seed(1234)
fit.km <- kmeans(wine, 3, nstart=25)
fit.km$size
```


#### Now we want to evaluate how well this clustering does.

####Exercise 5: using the table() function, show how the clusters in fit.km$clusters compares to the actual wine types in wine$Type. Would you consider this a good clustering?
```{r}
table(fit.km$cluster)
```

#### Exercise 6: Visualize these clusters using  function clusplot() from the cluster library Would you consider this a good clustering?

library(cluster)
clusplot(fit.km$cluster)
```



